### 2020-04-07 manuscript/manuscript.RmD
- major updates on theory, analyses, and osm

---
title             : "How Do Like and Dislike Buttons Affect Communication? A Privacy Calculus Approach to Understanding Self-Disclosure Online in a One-Week Field Experiment"
shorttitle        : "Popularity cues and the privacy calculus"
author: 
  - name          : "Author"
    affiliation   : "1"
    corresponding : yes
    address       : "Postal address"
    email         : "my@email.com"
affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
authornote: |
abstract: |
    Äccording to the privacy calculus, both privacy concerns and expected gratifications explain self-disclosure online. So far, little is known about whether the privacy calculus can be used to predict observations of actual authentic behavior, and whether the privacy calculus can be influenced by the design of online websites---for example, by implementing popularity cues such as like and dislike buttons. To answer this question, we ran a preregistered one-week field experiment, in which participants were randomly distributed to three different websites where they could discuss a current political topic. First, the results showed that privacy calculus variables predicted a considerable share of actual self-disclosure. Second, we found that the impact of implementing popularity cues was negligible. In conclusion, the results demonstrate that self-disclosure online can be explained by, for example, privacy concerns and psychological gratifications. This finding has several implications---for example, it provides further evidence against the privacy paradox.
keywords          : "privacy calculus, self-disclosure, popularity cues, structural equation modeling, preregistration"
bibliography      : 
  - [../bibliography/r-references.bib]
  - [../bibliography/references.bib]
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : yes
wordcount         : "`r wordcountaddin::word_count('../manuscript/manuscript.Rmd')`"
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
# output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
---

```{r include=F}
# knitr::opts_knit$set(root.dir = "../")
knitr::opts_chunk$set(cache=F)
```

@@

# load bib
r_citations <- cite_r("../bibliography/r-references.bib"
                      , footnote = FALSE
                      , withhold = FALSE
                      , pkgs = c("tidyverse", "lavaan", "papaja", "pwr", "quanteda", "semTools")
                      )
load("../data/workspace.RData")
```

@@

Rather than further pursuing the privacy paradox, a large share of current day research posits that self-disclosure online can be explained---at least partly---by means of the so-called _privacy-calculus_ [@krasnovaOnlineSocialNetworks2010]. The privacy calculus builds on the work of @lauferPrivacyConceptSocial1977 and claims that both expected risks *and* expected benefits explain self-disclosure. Specifically, by operationalizing expected risks as privacy concerns, several studies have shown that experiencing greater privacy concerns is related to disclosing less information [e.g., @heirmanPredictingAdolescentsDisclosure2013].

@@

Second, current research on the privacy calculus is often criticized for not explicitly focusing on the deliberation process of self-disclosure. According to critics [e.g., @knijnenburgDeathPrivacyCalculus2017], showing that concerns and gratifications both correlate with self-disclosure is not evidence for any substantial or explicit weighing of pros and cons. We agree and consider it necessary to now explicitly focus on the privacy deliberation process itself. Moreover, and on a more general level, we aim to gauge the usefulness of further extending the privacy calculus model by adding new variables such as privacy deliberation, trust, and self-disclosure self-efficacy.

@@

To test our research questions, we conducted a preregistered online field experiment, drawing from a representative sample of the German population. Participants were randomly distributed to one of three different websites, which either featured only a like button, both a like and a dislike button, or no buttons at all. Over the course of one week participants had the chance to discuss a topical issue (i.e., prevention of terrorist attacks in Germany). Afterward, they answered our follow-up questionnaire with items pertaining to the privacy calculus variables.

@@

Being a primary means of regulating privacy [e.g., @masurSituationalPrivacySelfdisclosure2018], self-disclosure is our key variable of interest. There are two different understandings of self-disclosure in the literature: The first defines self-disclosure as _deliberate_ acts of sharing truthful information about the self with others [@jourardTransparentSelf1964]. The second considers _all_ acts of sharing information---whether active or passive, deliberate or unwitting---as self-disclosure, because each piece of information shared allows meaningful inferences to be made about a person [e.g., @watzlawickPragmaticsHumanCommunication2011]. In this paper we follow the latter approach, not least because recent years have vividly illustrated how it is possible to derive a plethora of insights about a person simply by analyzing his or her written communication [e.g., @kosinskiPrivateTraitsAttributes2013]. Moreover, independently from which position one chooses to adopt, it is possible to differentiate the content of self-disclosure into three different dimensions: breadth (i.e., number of topics covered), depth (i.e., intimacy of topics covered), and length (i.e., quantity of disclosure) [e.g., @omarzuDisclosureDecisionModel2000]. In this study we mainly focus on communication quantity, as we consider communication quantity to be a necessary precondition and hence valid proxy for self-disclosure.

Privacy concerns have been defined as follows: "Concerns about online privacy represent how much an individual is motivated to focus on his or her control over a voluntary withdrawal from other people or societal institutions on the Internet, accompanied by an uneasy feeling that his or her privacy might be threatened" [AUTHOR]. Previous research has found that people who are more concerned about their privacy than others are less inclined to share personal information [e.g., @baruhOnlinePrivacyConcerns2017; @dienlinPrivacyParadoxRelic2015; @heirmanPredictingAdolescentsDisclosure2013].

@@

In the current literature on the privacy calculus there still seems to be a shortage of studies that explicitly analyze the process of actively comparing the pros and cons of disclosing information, even though this point of criticism has been levelled several times [e.g., @knijnenburgDeathPrivacyCalculus2017] and even though other fields such as behavioral economics have long focused on the underlying problem [e.g., @zhuPrivacyCalculusIts2017]. The criticism is justified, because showing that experiencing privacy concerns and expecting gratifications are related to self-disclosure does not necessarily imply that an explicit weighing took place. Hence, we argue that research on the privacy calculus would benefit significantly from addressing a novel concept that might best be termed _privacy deliberation_, which we define as the extent to which individual people explicitly compare positive and negative potential outcomes before communicating with others.

On the one hand, it seems plausible to suggest that deliberating about one's privacy would dampen subsequent self-disclosure, because refraining from regular communication---the primary means of connecting with others---requires at least a minimum of active and hence deliberate restraint. On the other hand, deliberating about one's privacy might also increase self-disclosure, as after having actively deliberated about the potential consequences, a person concerned about his or her privacy might arrive at the conclusion that in this situation self-disclosure is not only appropriate but expedient. In light of the paucity of studies and the plausibility of both effects, we formulate the following research question:

@@

Several attempts have already been made to expand the privacy calculus [e.g., @dinevExtendedPrivacyCalculus2006]. Additional variables such as privacy self-efficacy or trust have been introduced. Building on @dienlinExtendedPrivacyCalculus2016, self-efficacy in the context of the privacy calculus captures whether people believe in their own capability to implement particular privacy behaviors in the future that foster either self-withdrawal (e.g., deleting inappropriate content) or self-disclosure (e.g., publishing a blog post). Thus far, several studies have found that people who report more privacy self-efficacy also self-withdraw more online than others [e.g., @chenRevisitingPrivacyParadox2018].

Trust can be conceptualized in two different ways [@gefenTrustTAMOnline2003]: It either captures "specific beliefs dealing primarily with the integrity, benevolence, and ability of another party" [@gefenTrustTAMOnline2003, p. 55] or a "general belief that another party can be trusted" [@gefenTrustTAMOnline2003, p.55]. @gefenTrustTAMOnline2003 prioritize specific trust beliefs (p. 60). In the online context, it is important to differentiate among several targets of trust [@sollnerWhyDifferentTrust2016]. For example, one can differentiate between (a) the information system, (b) the provider, (c) the Internet, and (d) the community of other users [@sollnerWhyDifferentTrust2016]. Trust plays a key role in online communication [e.g., @metzgerPrivacyTrustDisclosure2004]. For example, it has been demonstrated that people who put more trust in the providers of networks also disclose more personal information [see, e.g., @liEmpiricalStudiesOnline2011].

@@

As the privacy calculus states that both benefits and costs determine behavior, we suggest that popularity cues such as like and dislike buttons, which are categorized as "paralinguistic digital affordances" [@carrPredictingThresholdPerceived2018, p. 142], perfectly epitomize benefits and costs. The like button is positive; it expresses an endorsement, a compliment, a reward [e.g., @sumnerFunctionalApproachFacebook2017]. However, communication online is also often characterized by negative and critical debates [e.g., @ziegeleDynamicsOnlineNews2017]. As the dislike button is a major means of downgrading content it represents the cost and risk factor of the privacy calculus well---in fact, its stark negative effect might also explain why only a handful of major websites have implemented a dislike button (e.g., reddit.com or stackexchange.com).

Popularity cues have been shown to impact behavior. For example, a large-scale field experiment in which 101,281 comments were analyzed found that comments with dislikes were more likely to receive further dislikes [@muchnikSocialInfluenceBias2013]. @stroudRecommendRespectAltering2017 demonstrated that when users had a different opinion than the one that was communicated in a post, they were more likely to click on a button labelled _respect_ compared to a button labelled _like_.

In this vein it seems plausible to suggest that popularity cues might also impact the privacy calculus. First, on a very primordial level, popularity cues should serve as a means of reward and punishment, as a mechanism of instrumental conditioning [@skinnerScienceHumanBehavior2014]. Specifically, we argue that being complimented with a like should encourage future self-disclosure, while being punished with a dislike should inhibit future disclosure. Similarly, like buttons should be associated with being able to garner positive feedback, so implementing a like-button---similar to a compliment in the offline world---might leverage gratifications. Implementing a like or a dislike button might also bring people to more actively deliberate about whether or not it is actually worthwhile to disclose information. If both like and dislike buttons are present, privacy deliberation should increase even further. Finally, because people who are more concerned about their privacy are also more shy and risk averse [@dienlinPsychologyPrivacyAnalyzing2017], implementation of the dislike button should both stir privacy concerns and stifle self-disclosure. For a simplified overview of our theoretical model, see Figure \@ref(fig:model).

@@

We preregistered the study's hypotheses, sample size, materials, analyses, and exclusion criteria (https://osf.io/a6tzc/?view_only=5d0ef9fe5e1745878cd1b19273cdf859 [^prereg]). We needed to change our pre-defined plan in some cases. For a full account of all changes, see online supplementary material (OSM), https://osf.io/hcqat/?view_only=5db35868738d40609b11e58cc343a9b0. New analyses that were not preregistered appear in the section on exploratory analyses. For example, we also measured additional variables that were not included in the preregistration (e.g., specific gratifications and general trust; see below), whose results we report as exploratory analyses.

[^prereg]: To find the actual preregistration on the website, click _View Registration Form_ (blue button on the right hand side).

@@

We collaborated with a professional panel agency to recruit participants. As an incentive, participants were awarded digital points, which they could use to get special offers from other companies. In order to be able to take part, participants needed to be above the age of 18 and had to live in Germany. In a first step, the agency sent their panel members an invitation to participate in the study (_invitation_). In this invitation, panel members were asked whether they would be interested in taking part in a study analyzing the threat posed by current terrorist attacks in Germany.[^1] If panel members decided to take part, they were subsequently sent the first questionnaire (_T1_) in which we asked about the participants’ sociodemographics, provided more details about the study, and included a registration link for the website. Afterward, the participants were randomly assigned to one of the three websites. After registration participants had the chance to discuss the topic of the terrorism threat in Germany over the course of one week (_field_). Afterward, participants received a follow-up questionnaire in which we collected the self-reported measures (_T2_). Measures were collected after and not before the field phase in order not to prime participants or reveal our primary research interest.

@@

We programmed an online website based on the open-source software discourse (https://www.discourse.org/). We conducted several pretests with students from the local university to make sure the website had an authentic feel (see Figure \@ref(fig:website)). Participants used the website actively: Overall, users spent `r sum(d$time_read/60) %>% round(digit = 0)` minutes online and wrote `r sum(d$post_count)` comments (for an example of communication that took place, see Figure \@ref(fig:comments)).

@@

We ran a priori power analyses to determine how many participants to recruit. The power analysis was based on the smallest effect size of interest [SESOI; @lakensEquivalenceTestingPsychological2018]. Thus, we defined an effect size that we would consider enough to support our hypotheses. Because small effects should be expected when researching aspects of privacy online [e.g., @baruhOnlinePrivacyConcerns2017], with small effects beginning at an effect size of _r_ = .10 [@cohenPowerPrimer1992], we set our SESOI to be r = `r r_sesoi`. Our aim was to be able to detect this SESOI with a probability of at least 95%. Using the regular alpha level of 5%, this leads to a minimum sample size of _n_ = `r n_desired`. In the end, we were able to include n = `r n_final` in our analyses (see below). This means that our study had a probability (power) of `r power_achieved * 100`% of finding an effect at least as large as _r_ = `r r_sesoi`. Put differently, we were able to make reliable inferences about effects at least as big as _r_ = `r r_sensitive`.

We collected a representative sample of the German population in terms of age, sex, and federal state. T1 was completed by _N_ = `r n_t1` participants and T2 by _n_ = `r n_t2` participants. We connected the data from T1, participants' behavior on the platform, and T2 by means of objective and automated processes (e.g., participants were matched via tokens or IP addresses). So far, we have been able to match the data for _n_ = `r n_final` participants. The matching process is not yet finished, and we will try to match the remaining cases manually. We excluded _n_ = `r n_speeding` participants who finished the questionnaire at T2 in less then three minutes, which we considered to be unreasonably fast. The sample characteristics at T1 were as follows: 
Age = `r round(age_t1_m, 0)` years, sex = `r round(male_t1_m, 2)*100`% male, college degree = `r round(college_t1_m, 2)*100`%.
The characteristics of the final sample were as follows: 
Age = `r round(age_final_m, 0)` years, sex = `r round(male_final_m, 2)*100`% male, college degree = `r round(college_final_m, 2)*100`%.
Hence, despite dropout, T2 can also be considered a largely representative sample of the German population.

@@

In what follows, we present the materials we used to measure our variables. To gauge the variables' factor validity, we ran confirmatory factor analyses (CFA). If the CFAs revealed insufficient fit, we deleted individual items. All items were formulated as statements to which participants indicated their (dis-)agreement on a bipolar 7-point scale. Answer options were as follows: -3 (_strongly disagree_), -2 (_disagree_), -1 (_slightly disagree_), 0 (_neutral_), +1 (_slightly agree_), +2 (_agree_), +3 (_strongly agree_). In the questionnaire, all items measuring a variable were presented on the same page in a randomized order.

For an overview of the means, standard deviations, factorial validity, and reliability, see Table \@ref(tab:CFA). For an overview of the variables' distributions, see Figure \@ref(fig:corrplot). For the exact wording of all items and their individual distributions, see the OSM.

@@

```{r corrplot, fig.cap="Above diagonal: zero-order correlation matrix; diagonal: density plots for each variable; below diagonal: bivariate scatter plots for zero-order correlations. Solid regression lines represent linear regressions, dotted regression lines represent quadratic regressions. Calculated with the model predicted values for each variable (baseline model).", fig.height=8, fig.width=8, out.width=".9\\textwidth", fig.align='center', fig.pos = "!h", warning=F}
print(plot_fs_cor)
```

@@

Privacy concerns were measured with seven items based on @buchananDevelopmentMeasuresOnline2007 One example item was "When using the participation platform, I had concerns about my privacy". One item had to be deleted due to poor psychometric properties.

@@

Next, we differentiated between two separate types of gratification. _General gratifications_ were measured with five items based on @sunLocationInformationDisclosure2015 One example item was "Using the participation platform has paid off for me". _Specific gratifications_ were measured with 15 items on five different subdimensions with three items each. The scaled was loosely based on @schererGratifikationMinuteZeitnahe2002 Example items were: "Using the participation platform made it possible for me to" ... "learn things I would not otherwise have noticed" (information), "react to a subject that is important to me" (relevance), "engage politically" (political participation), "try to improve society" (idealism), and "soothe my guilty consciences" (extrinsic benefits).

### Self-disclosure self-efficacy

Self-disclosure self-efficacy was captured with six self-designed items, which captured whether participants felt that they had sufficient self-efficacy to write a comment on the platform. For example, we asked "I felt technically competent enough to write a comment." Two items, which were inverted, had to be deleted due to poor psychometric properties.

Next, we differentiated between two separate types of trust. _General trust_ was operationalized based on @sollnerWhyDifferentTrust2016 for three targets (i.e., provider, website, and other users) with one item each. One example items was "The operators of the participation platform seemed trustworthy."
_Specific trust_ was operationalized for the same three targets with three subdimensions each (i.e., ability, benevolence/integrity, and reliability), which were measured with one item each. Example items were "The operators of the participation platform have done a good job" (ability), "The other users had good intentions" (benevolence/integrity), "The website worked well" (reliability). The results showed that the provider and website targets were not sufficiently distinctive, as was evidenced by the existence of a Heywood case. We hence adapted the scale to combine these two targets. The updated scale exhibited adequate fit.

@@

@@

Self-disclosure was calculated by taking the natural logarithm of the number of comments each participant posted.

>>>>

---
title             : "How Do Like and Dislike Buttons Affect Communication? A Privacy Calculus Approach to Understanding Self-Disclosure Online in a One-Week Field Experiment"
shorttitle        : "Popularity cues and the privacy calculus"
author: 
  - name          : "Author"
    affiliation   : "1"
    corresponding : yes
    address       : "Postal address"
    email         : "my@email.com"
affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
authornote: |
abstract: |
    According to the privacy calculus, both privacy concerns and expected gratifications explain self-disclosure online. So far, most findings are based on self-reports, and little is known about whether the privacy calculus can be used to explain observations of actual authentic behavior. Likewise, we still know comparatively little whether the privacy calculus can be influenced by the design of online websites, including for example popularity cues such as like and dislike buttons. To answer these questions, we ran a preregistered one-week field experiment. Participants were randomly distributed to three different websites where they could discuss a current political topic. The final sample consisted of `r knitr::load_cache(label = "load-data", object = "n_matched")` participants. The results showed that privacy calculus variables predicted a considerable share of actual self-disclosure. The impact of popularity cues was negligible. In conclusion, the results demonstrate that self-disclosure online can be explained by privacy concerns and specific gratifications. This finding has several implications. For example, it provides further evidence against the privacy paradox.
keywords          : "privacy calculus, self-disclosure, popularity cues, structural equation modeling, preregistration"
bibliography      : 
  - [../bibliography/r-references.bib]
  - [../bibliography/references.bib]
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : yes
wordcount         : "`r wordcountaddin::word_count('../manuscript/manuscript.Rmd')`"
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
# output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
---

```{r include=F}
knitr::opts_chunk$set(cache=F)
```

@@

# load bib
r_citations <- cite_r("../bibliography/r-references.bib"
                      , footnote = FALSE
                      , withhold = FALSE
                      , pkgs = c("tidyverse", "lavaan", "papaja", "pwr", "quanteda", "semTools")
                      )
```

```{r load-data, cache=T, include=F}
load("../data/workspace.rData")
```

@@

Rather than further pursuing the privacy paradox, a large share of current day research posits that self-disclosure online can be explained---at least partly---by means of the so-called _privacy-calculus_ [@krasnovaOnlineSocialNetworks2010]. The privacy calculus builds on the work of @lauferPrivacyConceptSocial1977 and claims that both expected risks *and* expected benefits explain self-disclosure. Specifically, by operationalizing expected risks as privacy concerns, several studies have shown that experiencing greater privacy concerns is related to disclosing less information [@heirmanPredictingAdolescentsDisclosure2013; koohikamaliInvestigationDynamicModel2019].

@@

Second, current research on the privacy calculus is often criticized for not explicitly focusing on the deliberation process of self-disclosure. 
According to critics [e.g., @knijnenburgDeathPrivacyCalculus2017], showing that concerns and gratifications both correlate with self-disclosure is not evidence for any substantial or explicit weighing of pros and cons.

We agree and consider it necessary to now explicitly focus on the privacy deliberation process itself. 
Moreover, and on a more general level, we aim to gauge the usefulness of further extending the privacy calculus model by adding new variables such as privacy deliberation, trust, and self-disclosure self-efficacy.

@@

To test our research questions, we conducted a preregistered online field experiment, drawing from a representative sample of the German population. Participants were randomly distributed to one of three different websites, which either included only a like button, both a like and a dislike button, or no buttons at all. Over the course of one week participants had the chance to discuss a topical issue (i.e., prevention of terrorist attacks in Germany). Afterward, they answered our follow-up questionnaire with items pertaining to the privacy calculus variables.

@@

Being a primary means of regulating privacy [e.g., @masurSituationalPrivacySelfdisclosure2018], self-disclosure is our key variable of interest. 
There are two different understandings of self-disclosure in the literature: The first defines self-disclosure as _deliberate_ acts of sharing truthful information about the self with others [@jourardTransparentSelf1964]. 
The second considers _all_ acts of sharing information---whether active or passive, deliberate or unwitting---as self-disclosure, because each piece of information shared allows meaningful inferences to be made about a person [e.g., @watzlawickPragmaticsHumanCommunication2011]. 
In this paper we follow the latter approach, not least because recent years have vividly illustrated how it is possible to derive a plethora of insights about a person simply by analyzing his or her written communication [e.g., @kosinskiPrivateTraitsAttributes2013]. 
Moreover, independent from which position one chooses to adopt, it is possible to differentiate the content of self-disclosure into three different dimensions: breadth (i.e., number of topics covered), depth (i.e., intimacy of topics covered), and length (i.e., quantity of disclosure) [e.g., @omarzuDisclosureDecisionModel2000]. 
In this study we mainly focus on communication quantity, as we consider communication quantity to be a necessary precondition and hence valid proxy for self-disclosure.

Privacy concerns have been defined as follows: 
"Concerns about online privacy represent how much an individual is motivated to focus on his or her control over a voluntary withdrawal from other people or societal institutions on the Internet, accompanied by an uneasy feeling that his or her privacy might be threatened" [AUTHOR]. 
Previous research has found that people who are more concerned about their privacy than others are less inclined to share personal information [@baruhOnlinePrivacyConcerns2017; @dienlinPrivacyParadoxRelic2015; @heirmanPredictingAdolescentsDisclosure2013; @koohikamaliInvestigationDynamicModel2019].

@@

In the current literature on the privacy calculus there still seems to be a shortage of studies that explicitly analyze the decision process of actively comparing the pros and cons of disclosing information, although this point of criticism has been leveled several times [e.g., @knijnenburgDeathPrivacyCalculus2017] and although other fields such as behavioral economics have long focused on the underlying problem [e.g., @zhuPrivacyCalculusIts2017]. 
This criticism is justified.
The observation that both experiencing privacy concerns and expecting gratifications are related to self-disclosure does not bit itself necessitate an explicit weighing process 
Hence, we argue that the research on the privacy calculus would benefit significantly from analyzing this decision process explicitly. 
Building on @omarzuDisclosureDecisionModel2000 and @altmanPrivacyConceptualAnalysis1976, we hence address a novel concept that might best be termed _privacy deliberation_, which we define as the extent to which individual people explicitly compare positive and negative potential outcomes before communicating with others.

On the one hand, it seems plausible that deliberating about one's privacy would dampen subsequent self-disclosure, because refraining from regular communication---the primary means of connecting with others---requires at least a minimum of active and hence deliberate restraint. 
On the other hand, deliberating about one's privacy might also increase self-disclosure, as after having actively deliberated about the potential consequences, a person concerned about his or her privacy might arrive at the conclusion that in this situation self-disclosure is not only appropriate but expedient. 
In light of the paucity of empirical studies and the plausibility of both effects, we formulate the following research question:

@@

Several attempts have already been made to expand the privacy calculus [e.g., @dinevExtendedPrivacyCalculus2006]. 
Additional variables such as self-efficacy or trust have been introduced. 
Self-efficacy in the context of the privacy calculus captures whether people believe in their own capability to implement particular privacy behaviors in the future [@dienlinExtendedPrivacyCalculus2016].
These privacy behaviors can either refer to self-withdrawal (e.g., deleting inappropriate content) or self-disclosure (e.g., publishing a blog post).
Thus far, several studies have found that people who report more privacy self-efficacy also self-withdraw more online than others [e.g., @chenRevisitingPrivacyParadox2018].
In light of our focus on self-disclosure, in this study we investigate the influence of self-disclosure self-efficacy.

Trust can be conceptualized in two different ways [@gefenTrustTAMOnline2003]: It either captures "_specific_ beliefs dealing primarily with the integrity, benevolence, and ability of another party" [@gefenTrustTAMOnline2003, p. 55, emphasis added] or a "_general_ belief that another party can be trusted" [@gefenTrustTAMOnline2003, p.55, emphasis added]. 
Whereas specific trust beliefs focus on the causes of trust, general trust beliefs focus on the experience of trust.
@gefenTrustTAMOnline2003 prioritize specific trust beliefs (p. 60). 
In the online context, it is important to differentiate among several targets of trust [@sollnerWhyDifferentTrust2016]. 
Potential targets include (a) the information system, (b) the provider, (c) the Internet, and (d) the community of other users [@sollnerWhyDifferentTrust2016]. 
Trust plays a key role in online communication [@metzgerPrivacyTrustDisclosure2004]. 
For example, it has been demonstrated that people who put more trust in the providers of networks also disclose more personal information [@liEmpiricalStudiesOnline2011].

@@

As the privacy calculus states that both benefits and costs determine behavior, we suggest that popularity cues such as like and dislike buttons, which are categorized as "paralinguistic digital affordances" [@carrPredictingThresholdPerceived2018, p. 142], perfectly epitomize benefits and costs. 
The like button is positive; it expresses an endorsement, a compliment, a reward [@carrPredictingThresholdPerceived2018; @sumnerFunctionalApproachFacebook2017]. 
However, communication online is also often characterized by negative and critical debates [@ziegeleDynamicsOnlineNews2017]. 
As the dislike button is a major means of downgrading content it represents the cost and risk factor of the privacy calculus well. 
In fact, its stark negative effect might also explain why to date only a handful of major websites have implemented it (e.g., youtube, reddit or stackexchange).

Paralinguistic digital affordances and/or popularity cues have been shown to impact behavior [@kramerMasteringChallengeBalancing2020; @treptePrivacyCalculusContextualized2020].
For example, a large-scale field experiment in which 101,281 comments were analyzed found that comments with dislikes were more likely to receive further dislikes [@muchnikSocialInfluenceBias2013]. 
@stroudRecommendRespectAltering2017 demonstrated that when users had a different opinion than the one that was communicated in a post, they were more likely to click on a button labelled _respect_ compared to a button labelled _like_.

In this vein it seems plausible that popularity cues might also impact the privacy calculus [kramerMasteringChallengeBalancing2020].
First, on a primordial level, popularity cues serve as a means of reward and punishment, affecting behavior via instrumental conditioning [@skinnerScienceHumanBehavior2014]. 
Specifically, being complimented with a like should encourage future self-disclosure, while being punished with a dislike should inhibit future disclosure. 
Similarly, like buttons should be associated with being able to garner positive feedback, so implementing a like-button---similar to a compliment in the offline world---might leverage gratifications. Implementing a like or a dislike button might also bring people to more actively deliberate about whether or not it is actually worthwhile to disclose information. 
If both like and dislike buttons are present, privacy deliberation should increase even further. 
Finally, because people who are more concerned about their privacy are also more shy and risk averse [@dienlinPsychologyPrivacyAnalyzing2017], implementation of the dislike button should both stir privacy concerns and stifle self-disclosure. 
For a simplified overview of our theoretical model, see Figure \@ref(fig:model).

@@

The online supplementary material (OSM) of this study include the data, research material, analyses scripts, and a reproducible version of this manuscript (see https://osf.io/hcqat/?view_only=5db35868738d40609b11e58cc343a9b0)
We preregistered the study using the registration form _OSF Prereg_, which includes hypotheses, sample size, materials, analyses, and exclusion criteria (see https://osf.io/a6tzc/?view_only=5d0ef9fe5e1745878cd1b19273cdf859). 
We needed to change our pre-defined plan in some cases. 
For a full account of all changes, see OSM. 
New analyses that were not preregistered appear in the section on exploratory analyses. 
For example, we also measured two additional variables that were not included in the preregistration (e.g., _specific_ gratifications and _general_ trust; see below), which are included in the exploratory analyses.

@@

We collaborated with a professional panel agency to recruit participants. 
As incentive, participants were awarded digital points, which they could use to get special offers from other companies. 
Participants were above the age of 18 and lived in Germany. 
In a first step, the agency sent their panel members an invitation to participate in the study (_invitation_). 
In this invitation, panel members were asked to participate in a study analyzing the current threat posed by terrorist attacks in Germany.[^1] 
Members who decided to take part were subsequently sent the first questionnaire (_T1_), in which we asked about their sociodemographics, provided more details about the study, and included a registration link for the website. 
Afterward, participants were randomly assigned to one of the three websites. 
After registration participants had the chance to discuss the topic of the terrorism threat in Germany over the course of one week (_field_). 
Subsequently, participants received a follow-up questionnaire in which we collected the self-reported measures (_T2_). 
Measures were collected after and not before the field phase in order not to prime participants or reveal our primary research interest.

@@

We programmed an online website based on the open-source software discourse (https://www.discourse.org/). 
We conducted several pretests with students from the local university to make sure the website had an authentic feel (see Figure \@ref(fig:website)). 
Participants used the website actively: Overall, they spent `r sum(d_raw$time_read, na.rm = TRUE) %>% "/"(60) %>% round(0) %>% format(big.mark = ",")` minutes online, wrote `r sum(d_raw$post_count, na.rm = TRUE) %>% format(big.mark = ",")` comments, and left `r sum(d_raw$reactions, na.rm = TRUE)` popularity cues. 
For an example of communication that took place, see Figure \@ref(fig:comments).

@@

We ran a priori power analyses to determine how many participants to recruit. The power analysis was based on the smallest effect size of interest [SESOI; @lakensEquivalenceTestingPsychological2018]. 
Thus, we defined an effect size that we would consider enough to support our hypotheses. 
Because small effects should be expected when researching aspects of privacy online [e.g., @baruhOnlinePrivacyConcerns2017], with small effects beginning at an effect size of _r_ = .10 [@cohenPowerPrimer1992], we set our SESOI to be r = `r r_sesoi %>% my_round("std")`. 
Our aim was to be able to detect this SESOI with a probability of at least 95%. Using the regular alpha level of 5%, this leads to a minimum sample size of _n_ = `r n_desired %>% format(big.mark = ",")`. 
In the end, we were able to include n = `r n_final` in our analyses (see below). 
This means that our study had a probability (power) of `r power_achieved * 100`% of finding an effect at least as large as _r_ = `r r_sesoi %>% my_round("std")`. 
Put differently, we were able to make reliable inferences about effects at least as big as _r_ = `r r_sensitive %>% my_round("std")`.

We collected a representative sample of the German population in terms of age, sex, and federal state. 
`r n_t1 %>% format(big.mark = ",")` participants completed the survey at T1, `r n_users` participants created a user account on the website,  and `r n_t2` participants completed the survey at T2. 
Using tokens and IP addresses, we connected the data from T1, participants' behavior on the platform, and T2 by means of objective and automated processes.
The data for _n_ = `r n_matched` participants could be matched successfully across all three platforms. 
We excluded _n_ = `r n_speeding` participants who finished the questionnaire at T2 in less then three minutes, which we considered to be unreasonably fast. 
To detect corrupt data, we calculated Cook's distance. 
We excluded `r n_resp_pattern` participants because they provided clear response patterns. 
The final sample included `r n_final` participants.
The sample characteristics at T1 and T2 were as follows: 
T1: Age = `r round(age_t1_m, 0)` years, sex = `r round(male_t1_m, 2)*100`% male, college degree = `r round(college_t1_m, 2)*100`%.
T2: Age = `r round(age_final_m, 0)` years, sex = `r round(male_final_m, 2)*100`% male, college degree = `r round(college_final_m, 2)*100`%.
(One participant did not report his or her sex.) 
<!-- Despite the dropout and although somewhat higher-educated, also T2 can hence be considered a largely representative sample of the German population. -->

@@

In what follows, we present the materials we used to measure our variables. 
Wherever possible, we operationalized our variables using established measures.
Where impossible (for example, to date there exists no scale on privacy deliberation), we self-designed novel items that were pretested in terms of legibility and/or understandability.
To gauge the variables' factor validity, we ran confirmatory factor analyses (CFA). 
If the CFAs revealed insufficient fit, we deleted individual items. 
All items were formulated as statements to which participants indicated their (dis-)agreement on a bipolar 7-point scale. 
Answer options were as follows: -3 (_strongly disagree_), -2 (_disagree_), -1 (_slightly disagree_), 0 (_neutral_), +1 (_slightly agree_), +2 (_agree_), +3 (_strongly agree_). 
In the questionnaire, all items measuring a variable were presented on the same page in a randomized order.

For an overview of the means, standard deviations, factorial validity, and reliability, see Table \@ref(tab:CFA). 
For an overview of the variables' distributions, see Figure \@ref(fig:corrplot). 
For the exact wording of all items and their individual distributions, see OSM.

@@

```{r corrplot, fig.cap="Above diagonal: zero-order correlation matrix; diagonal: density plots for each variable; below diagonal: bivariate scatter plots for zero-order correlations. Solid regression lines represent linear regressions, dotted regression lines represent quadratic regressions. Calculated with the model predicted values for each variable (baseline model).", fig.height=8, fig.width=8, out.width=".9\\textwidth", fig.align='center', fig.pos = "!h", warning=F}
print(fig_fs_cor)
```

@@

Privacy concerns were measured with seven items based on @buchananDevelopmentMeasuresOnline2007.
One example item was "When using the participation platform, I had concerns about my privacy". 
One item had to be deleted due to poor psychometric properties.

@@

Next, we differentiated between two separate types of gratification. 
_General gratifications_ were measured with five items based on @sunLocationInformationDisclosure2015.
One example item was "Using the participation platform has paid off for me". 
_Specific gratifications_ were measured with 15 items on five different subdimensions with three items each. 
The scaled was loosely based on @schererGratifikationMinuteZeitnahe2002.
Example items were: "Using the participation platform made it possible for me to" ... "learn things I would not otherwise have noticed" (information), "react to a subject that is important to me" (relevance), "engage politically" (political participation), "try to improve society" (idealism), and "soothe my guilty consciences" (extrinsic benefits).

@@

### Self-efficacy

Self-efficacy was captured with six self-designed items, which captured whether participants felt that they had sufficient self-efficacy to write a comment on the platform. 
For example, we asked "I felt technically competent enough to write a comment." 
Two items, which were inverted, had to be deleted due to poor psychometric properties.

Next, we differentiated between two separate types of trust. 
_General trust_ was operationalized based on @sollnerWhyDifferentTrust2016 for three targets (i.e., provider, website, and other users), with one item each. 
One example items was "The operators of the participation platform seemed trustworthy."
_Specific trust_ was operationalized for the same three targets with three subdimensions each (i.e., ability, benevolence/integrity, and reliability), which were measured with one item each. 
Example items were "The operators of the participation platform have done a good job" (ability), "The other users had good intentions" (benevolence/integrity), "The website worked well" (reliability). 
The results showed that the provider and website targets were not sufficiently distinctive, as was evidenced by the existence of a Heywood case. 
We hence adapted the scale to combine these two targets. 
The updated scale exhibited adequate fit.

@@

Self-disclosure was calculated by taking the log scale of the number of words each participant wrote in a comment plus the number of likes and dislikes, with likes and dislikes being multiplied by two. 
Like and dislike buttons were multiplied by two because, rudimentarily, like buttons abbreviate the sentence “I like” and dislike buttons the sentence “I dislike”.
The sum of words and likes/likes was log-scaled because the relative amount of self-disclosure diminishes the more a person has already said.
